\documentclass[11pt]{article}
\usepackage[a4paper,margin=1in]{geometry}
\usepackage{amsmath,amssymb}
\usepackage{booktabs}
\usepackage{hyperref}

\title{Q-FW (Yurtsever et al., 2022): Sections 3--4 Extraction and Comparison}
\author{}
\date{2026-03-09}

\begin{document}
\maketitle

\section{Scope}
This note summarizes the method in Sections 3 and 4 of:
\emph{Q-FW: A Hybrid Classical-Quantum Frank-Wolfe for Quadratic Binary Optimization}.
It then compares that method against the current implementation in this repository.

\section{Section 3: Problem Formulation (Extracted Method)}
The paper starts from a constrained quadratic binary optimization (QBO):
\begin{equation}
\min_{x \in \{0,1\}^n} \; x^\top Q x + 2 s^\top x
\quad \text{s.t.} \quad a_i^\top x = b_i,\; i=1,\dots,m.
\end{equation}

Main reformulation steps:
\begin{enumerate}
\item Symmetrize $Q$ without changing the objective.
\item Encode linear term into quadratic form for binary $x$.
\item Lift to matrix variable $X = xx^\top$ (rank-1, completely positive).
\item Rewrite objective and constraints linearly in $(x, X)$:
\begin{itemize}
\item $x^\top Q x = \mathrm{Tr}(QX)$,
\item $(a_i^\top x)^2 = b_i^2 \iff \mathrm{Tr}(A_i X) = b_i^2$, where $A_i = a_i a_i^\top$.
\end{itemize}
\item Replace nonconvex constraints with complete positivity constraints using the lifted matrix
\(W = \begin{bmatrix}1 & x^\top \\ x & X\end{bmatrix} \in \Delta^p\), with $p=n+1$.
\item Compact form:
\begin{equation}
\min_{W \in \Delta^p} \; \mathrm{Tr}(CW)
\quad \text{s.t.} \quad \mathcal{A}W = v,
\end{equation}
where $\mathcal{A}$ and $v$ collect all affine constraints, including
$\{a_i^\top x=b_i\}$, $\{\mathrm{Tr}(A_iX)=b_i^2\}$, $\mathrm{diag}(X)=x$, and $W_{1,1}=1$.
\end{enumerate}

This CP model is convex but NP-hard because of complete positivity.

\section{Section 4: Quantum Frank-Wolfe (FWAL/FWQP)}
The algorithm solves the CP model via an augmented Lagrangian:
\begin{equation}
\mathcal{L}_{\beta}(W; y) = \mathrm{Tr}(CW) + y^\top(\mathcal{A}W-v) + \frac{\beta}{2}\|\mathcal{A}W-v\|_2^2,
\end{equation}
with primal variable $W \in \Delta^p$, dual variable $y \in \mathbb{R}^d$, and penalty $\beta>0$.

At iteration $t$ (paper notation):
\begin{enumerate}
\item Increase penalty: $\beta_t = \beta_0\sqrt{t+1}$.
\item Primal gradient matrix:
\begin{equation}
G_t = C + \mathcal{A}^\ast y_t + \beta_t\,\mathcal{A}^\ast(\mathcal{A}W_t - v).
\end{equation}
\item LMO over complete-positive domain:
\begin{equation}
H_t \in \arg\min_{W\in\Delta^p} \mathrm{Tr}(G_tW).
\end{equation}
This is reduced to QUBO by solving
\begin{equation}
w_t \in \arg\min_{w\in\{0,1\}^p} w^\top G_t w,
\quad H_t = w_t w_t^\top.
\end{equation}
\item FW primal update:
\begin{equation}
W_{t+1} = (1-\eta_t)W_t + \eta_t H_t,
\quad \eta_t = \frac{2}{t+1}.
\end{equation}
\item Dual ascent:
\begin{equation}
g_t = \mathcal{A}W_{t+1} - v,
\quad y_{t+1} = y_t + \gamma_t g_t.
\end{equation}
In practice they use constant $\gamma_t=\beta_0$.
\item FWQP is recovered by disabling dual updates ($y_0=0$, $\gamma_t=0$).
\end{enumerate}

Rounding in Section 4:
\begin{enumerate}
\item From approximate $\hat W$, remove first row/column to get $\hat X$.
\item Compute best rank-1 approximation $\hat x\hat x^\top$ of $\hat X$ (Frobenius norm).
\item Optionally project $\hat x$ to feasible set of the original QBO (for permutations, Hungarian projection).
\end{enumerate}

\section{Differences vs Current Implementation}
\subsection{Summary}
The current code is a standard constrained FW-like loop in variable $x\in[0,1]^n$.
The paper's Q-FW is a CP-space FWAL/FWQP method in lifted variable $W\in\Delta^p$
with augmented-Lagrangian primal-dual updates.
This is a structural mismatch, not a minor hyperparameter mismatch.

\subsection{Concrete Mismatches}
\begin{table}[h]
\centering
\small
\begin{tabular}{p{0.23\linewidth} p{0.34\linewidth} p{0.35\linewidth}}
\toprule
Paper (Sections 3--4) & Current implementation & Impact \\
\midrule
Optimize lifted CP variable $W$ with constraints $\mathcal{A}W=v$. & Optimizes vector $x$ directly using $Q,c$ from original QBO; see \texttt{qfw\_solver.py} lines 121--123 and 170--194, and \texttt{gradient.py} lines 23--54. & Not the CP reformulation the paper proves/uses. \\

Use augmented Lagrangian $\mathcal{L}_\beta(W;y)$ with dual $y$ and increasing $\beta_t$. & No dual variable, no $\beta_t$, no primal-dual saddle updates in \texttt{qfw\_solver.py}. & Missing core FWAL/FWQP mechanism. \\

Primal gradient is matrix $G_t = C + \mathcal{A}^\ast y_t + \beta_t\mathcal{A}^\ast(\mathcal{A}W_t-v)$. & Uses $\nabla f(x)=(Q+Q^\top)x+c$; \texttt{gradient.py} lines 57--77. & Different objective geometry and search direction. \\

LMO solves $\min_{W\in\Delta^p}\mathrm{Tr}(G_tW)$ via binary QUBO in $w\in\{0,1\}^p$, then $H_t=ww^\top$. & LMO minimizes linear form $\nabla f(x)^\top s$ over sampled feasible bitstrings; \texttt{linear\_minimisation.py} lines 10--14 and 142--152. & Not equivalent to the paper's CP-domain LMO. \\

Dual step $y_{t+1}=y_t+\gamma_t(\mathcal{A}W_{t+1}-v)$. & Absent from \texttt{qfw\_solver.py}. & Cannot reproduce FWAL behavior/convergence statement in Prop. 1. \\

Rounding from $\hat W\to\hat X\to\hat x$ using best rank-1 approx and optional projection. & Threshold/greedy/randomized rounding on $x$; \texttt{rounding.py} lines 26--129. & Different decoding path than paper. \\

Constraint handling is encoded in affine map $\mathcal{A}$ and residual norm $\|\mathcal{A}W-v\|$. & Constraint treatment mostly by feasible sample filtering and heuristic rounding; also fallback may pick infeasible sample, \texttt{linear\_minimisation.py} lines 154--159. & Feasibility control differs from method definition. \\
\bottomrule
\end{tabular}
\end{table}

\section{What To Re-Implement Later (Correct Direction)}
For a faithful implementation of Sections 3--4, the future solver should:
\begin{enumerate}
\item Construct CP-lifted data $(C, \mathcal{A}, v)$ from the QBO instance.
\item Maintain state $(W_t, y_t, \beta_t)$.
\item Compute matrix gradient $G_t$ of the augmented Lagrangian.
\item Solve LMO in lifted domain (QUBO in $w$, then $H_t=w w^\top$).
\item Apply primal update for $W$ and dual ascent for $y$.
\item Decode $\hat W$ with rank-1 approximation and projection to original feasible set.
\end{enumerate}

\section{Code References}
The comparison above used these files:
\begin{itemize}
\item \texttt{quantum\_optimization/pipeline/solvers/frank\_wolfe/qfw\_solver.py}
\item \texttt{quantum\_optimization/pipeline/solvers/frank\_wolfe/gradient.py}
\item \texttt{quantum\_optimization/pipeline/solvers/frank\_wolfe/linear\_minimisation.py}
\item \texttt{quantum\_optimization/pipeline/solvers/frank\_wolfe/rounding.py}
\item \texttt{quantum\_optimization/pipeline/solvers/frank\_wolfe/qfw\_utils.py}
\end{itemize}

\end{document}
